---
title: "Reading Infographic"
author: "Amit Goldberg"
date: "2025-06-05"
format: html
---


```{r setup}
# Load required libraries
library(tidyverse)
library(here)
library(ggplot2)

# Read in dataset
reading_data <- read_csv(here("data", "updated_reading_dataset.csv"))
```


### Plot 1: Mood by Environment (Boxplot + Jitter)
```{r boxplot-jitter}
## Plot 1: Mood by Reading Environment
ggplot(reading_data, aes(x = Environment, y = Mood, fill = Environment)) +
  geom_boxplot(show.legend = FALSE, alpha = 0.6, outlier.shape = NA) +
  geom_jitter(width = 0.2, color = "black", alpha = 0.5, size = 2) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Mood While Reading by Environment",
    subtitle = "Quieter settings correlate with higher mood",
    x = "Reading Environment",
    y = "Mood Rating (1 = low, 5 = high)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 12)
  )
```
Mood ratings were highest in quiet environments like libraries and quiet cafes. Noisier spaces tended to lower my reading mood.


### Plot 2: Pages Read by Environment and Time of Day (Grouped Bar Plot)
```{r bar-plot}
## Plot 2: Pages Read by Environment and Time of Day

# Summarize total pages read by environment and time of day
pages_by_env_time <- reading_data %>%
  group_by(Environment, `Time of Day`) %>%
  summarize(Total_Pages = sum(`Pages Read`, na.rm = TRUE), .groups = "drop")

# Create grouped bar plot
ggplot(pages_by_env_time, aes(x = Environment, y = Total_Pages, fill = `Time of Day`)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_brewer(palette = "Pastel1") +
  labs(
    title = "When and Where I Read the Most",
    subtitle = "Grouped by Environment and Time of Day",
    x = "Reading Environment",
    y = "Total Pages Read",
    fill = "Time of Day"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 12)
  )
```
I read the most in libraries during the afternoon. Quiet settings and daytime hours support better focus and page count.


### Plot 3: Pages Read vs Duration of Time (Scatter Plot)
```{r scatter-plot}
## Plot 3: Reading Efficiency – Pages Read vs Duration

# Color points by reading environment to show differences in efficiency across settings
ggplot(reading_data, aes(x = `Duration (min)`, y = `Pages Read`, color = Environment)) +
  
  # Plot individual reading sessions as points
  geom_point(size = 3, alpha = 0.7) +
  
  # Add a dashed trend line to show general correlation (linear model)
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed", color = "gray30") +
  
  # Add titles and axis labels
  labs(
    title = "Does Reading Longer Mean Reading More?",
    subtitle = "Pages Read vs Duration (by Environment)",
    x = "Duration (minutes)",           # X-axis = session length
    y = "Pages Read",                   # Y-axis = pages read in that session
    color = "Environment"              # Legend label
  ) +
  
  # Use a clean, minimal theme
  theme_minimal(base_size = 13) +
  
  # Customize font sizes for readability
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 12)
  )
```
Longer sessions usually led to more pages read, but some environments helped me read more efficiently — finishing more pages in less time.

```{r setup for plot 4}
library(dplyr)   # For data wrangling
```

### Plot 4: Mood Over Time (Tile Plot)
```{r tile-plot}
## Plot 4: Tile Plot – Mood Over Time

# Ensure Date column is in Date format
reading_data <- reading_data %>%
  mutate(Date = as.Date(Date)) %>%
  arrange(Date)

# Create a tile-style mood strip plot
ggplot(reading_data, aes(x = Date, y = 1, fill = Mood)) +
  geom_tile(color = "white", height = 1) +
  scale_fill_gradientn(
    colors = c("#4575b4", "#91bfdb", "#e0f3f8", "#fee090", "#fc8d59", "#d73027"),
    limits = c(1, 5),
    name = "Mood"
  ) +
  labs(
    title = "How My Mood Shifted Over Time While Reading",
    subtitle = "Each tile represents one day; colors reflect mood",
    x = "Date",
    y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 12)
  )
```
This mood strip shows how my reading mood changed over time. Streaks of cool or warm tones reveal periods of high or low mood while reading.


### Write-Up

#### What patterns are you highlighting in each visualization? Why?

In the boxplot (Plot 1), I wanted to highlight how my mood varied across reading environments. It felt worth exploring because I noticed I tend to feel better in certain settings and this made it really clear. The grouped bar chart (Plot 2) let me compare how often I read in each environment and what the average mood was. I liked this one because it helped balance quantity with quality. The scatter plot (Plot 3) looked at reading efficiency — do longer sessions actually lead to more pages? Turns out, yes, but some spaces help me stay focused more than others. Finally, I created a mood strip (Plot 4) inspired by Ellie Highwood's temperature blanket. It gave me a cool way to see how my mood shifted over time. I could spot streaks of good or not-so-good days visually.

#### Outside of the data visualization, what aesthetic choices (e.g. color, font, arrangement) did you make and why?

I kept things minimal and clean so the focus stayed on the patterns in the data. I used color to carry the emotion, especially in the strip plot, where it helps you feel the changes over time. I also used straightforward titles and added some spacing between visuals in the infographic to keep it from feeling cluttered. Everything was arranged in a way that felt easy to follow from one plot to the next.

#### For each visualization, describe what examples inspired that visualization (cite the author)

Plot 1 was based on a jittered boxplot example we saw in class, and I liked how it showed both the range of values and individual points. For Plot 2, I followed a typical grouped bar chart format and tried to balance clarity and color to compare frequency and mood across environments. Plot 3 is a classic scatter plot with a smoothing line — I wanted it to show the relationship between reading time and pages read, and I used examples from online ggplot guides to figure out the styling. Plot 4 was inspired by temperature blanket visuals I’ve seen before, especially Ellie Highwood’s climate change piece. I liked the idea of showing mood over time through color, and wanted to try something visually similar for my own data.

#### Describe your coding process

I started by cleaning up the data. I made sure the date column was in the right format and double-checked for any missing values. I also summarized mood, pages read, and duration by environment to make certain plots easier to work with. For the first few visualizations, I stuck with geoms we used in class like `geom_boxplot`, `geom_jitter`, and `geom_col`. Once I felt more confident, I experimented with new ones like `geom_tile` for the strip plot. I also tried a radar chart using the `fmsb` package but ended up replacing it because it didn’t really communicate anything useful. Throughout the process, I adjusted color scales, added annotations, and tweaked themes to improve clarity. I committed my progress regularly to GitHub, which helped me stay organized and keep track of what was working (and what wasn’t).

#### Tools used

I used R and packages like ggplot2, dplyr, and fmsb (briefly). I used Canva to build the final infographic and GitHub for version control. I looked at examples from TidyTuesday, R Graph Gallery, and some of the code we worked on in class. I also used ChatGPT a couple times to refine code, troubleshoot weird errors, and pick color scales.






